---
title: "Joining Assignment"
author: "Giovanna Romariz Lino"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

```

# Load the datasets

```{r}

counties_demographics <- readRDS(here("data", "counties_demographics.rds"))
counties_elexresults <- readRDS(here("data", "counties_elexresults.rds"))
counties_rural_urban_suburban <- readRDS(here("data", "counties_rural_urban_suburban.rds"))



```

# Exercises

1)  

Join together the election results and urban/suburban/rural tables above, and then use the merged table to:

-   filter for just Virginia

-   group_by and summarize to calculate the mean pct of Trump 2020 vote (`pct_r_2020`) by urban/rural categories (`rural_urban`)

```{r}

inner_join(counties_elexresults, counties_rural_urban_suburban) %>% 
  filter (state_code == "VA") %>% 
  group_by(rural_urban) %>% 
  summarise(average = mean(pct_r_2020))
  


```

2)  

Using the joined election results / urban/suburban/rural table created above:

-   filter for just Virginia

-   count up how many counties were won by party in 2020 (`winner_2020`) by urban/rural categories

```{r}

inner_join(counties_elexresults, counties_rural_urban_suburban) %>% 
  filter (state_code == "VA") %>% 
  group_by(rural_urban, winner_2020) %>% 
  summarise(n())


```

3)  

Join together the election results and demographics tables above, and then use the joined table to:

-   filter for just Florida

-   group/summarize to calculate the mean pct of Trump 2020 vote in the counties above or below national average for college graduates (`pct_ed_college_all_abovebelow_natl`)

```{r}

inner_join(counties_elexresults, counties_demographics) %>% 
  filter (state_code == "FL") %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise(average = mean(pct_r_2020))



```

4)  

Using the joined table you made above:

-   group/summarize to calculate the mean pct of Trump 2020 vote by counties above or below the national average for college graduates (`pct_ed_college_all_abovebelow_natl`).

    Note we're looking for the entire country here.

```{r}

inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_ed_college_all_abovebelow_natl) %>% 
  summarise(average = mean(pct_r_2020))


```

Now using the joined datasets above, come up with two groupings you find interesting or newsworthy in a particular state, and replicate the steps above them them instead.

5)  By joining both datasets, I am grouping/summarizing to calculate the mean pct of Trump 2020 vote by counties above or below the national average for people who identify as nonwhite.  

```{r}

inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(pct_race_nonwhite_abovebelow_natl) %>% 
  summarise(average = mean(pct_r_2020))


```

6)  After joining both datasets, I want to find the amount of wins in 2020 for each party in each state. 
#This set of codes allows me to compare the amount of democratic vs republican winners in each state and which state had a major republican vs democratic victory. 

```{r}

inner_join(counties_elexresults, counties_demographics) %>% 
  group_by(state_name, winner_2020) %>% 
  summarise(n())

```

7)  Extra credit

Up until now we've joined two tables together. But we actually have three tables in this project. How might we join all three together into a single table, so all the measures for each county are combined in one place? (Hint: you can do this is multiple steps, doesn't have to be one step only.)


```{r}

mergedtables <- full_join(counties_elexresults, counties_demographics) 

full_join(mergedtables, counties_rural_urban_suburban, by = NULL, copy = FALSE)

#Here, I have assigned a new name for the two first tables I full joined, and then I full joined again the new assigned merged table to the last data to create all three tavles into a single one, with all the measures for each country combined in one place. 
  

```
